When to ignore perfect information: how iterative learning can lead to worse decisions

### 2021-03-26 manuscript/appendix.Rmd
:pen:

---
title: 'Appendix for: "Bad Forecast, Good Decision"'
date: "`r Sys.Date()`"
author: Carl Boettiger
journal: Appendix
layout: 3p
bibliography: refs.bib
#output:
#  tufte::tufte_handout:
#    dev: cairo_pdf
#    latex_engine: xelatex
output: 
  hrbrthemes::ipsum_pdf:
    dev: cairo_pdf
    latex_engine: xelatex

@@

This appendix provides a fully reproducible walk-through of the implementation of the analyses considered in this paper.  None of the methods applied here: Stochastic Dynamic Programming, Iterative Forecasts assessed with Proper Scoring Rules, or Adaptive Management, are novel

This appendix includes computer code in the R language, which can also be found in the corresponding R-Markdown document in the paper's GitHub repository, <https://github.com/cboettig/bad-forecast-good-decision>.  While computer code can be more verbose and difficult to read than mathematical formulae presented here, it is also less ambiguous.

The question of optimal harvest given a population growth model can be posed as a Markov Decision Process (MDP). @Marescot2013 provides an accessible overview of MDP problems and their solutions in conservation context.  @Reed1979 provides a formal proof of optimal harvest strategy for stochastic dynamics under the assumption of any concave growth function. The problem can be summarized as follows:

@@

This problem is already well studied, and it is worth noting that even under the simple utility function which assigns no value to fish that are not harvested, the optimal strategy still seeks to sustain the fish population indefinitely; as @Clark1973 shows for the deterministic function $f$ and @Reed1979 extended to the stochastic case.  As @Reed1979 proves, the optimal strategy can be characterized by a policy of "constant escapement," in which the manager adjusts the harvest effort each year in an effort to ensure the same number or biomass of fish are left in the sea each year.

The optimization problem can then be solved by dynamic programming, using the recursive Bellman equation as,

\begin{equation}
\end{equation}

This technique, known as stochastic dynamic programming (SDP) is widely used in conservation decision problems [e.g. @Mangel1985, @Marescot2013].

@@

@@

Passive adaptive management using a Bayesian learning scheme still learns the wrong model.

@@

```{r}
am <- am1$df %>% mutate(belief = am1$posterior$V2)
write_csv(am, "../data/am.csv")

>>>>

---
title: 'Appendix A'
date: "`r Sys.Date()`"
author: Carl Boettiger
journal: Appendix
layout: 3p
bibliography: refs.bib
#output:
#  tufte::tufte_handout:
#    dev: cairo_pdf
#    latex_engine: xelatex
output: 
  hrbrthemes::ipsum_pdf:
    dev: cairo_pdf
    latex_engine: xelatex

@@

This appendix provides a fully reproducible walk-through of the implementation of the analyses considered in this paper.  Because the methods applied here: Stochastic Dynamic Programming, Iterative Forecasts assessed with Proper Scoring Rules, or Adaptive Management, are not themselves novel, the reader is referred to previous reviews and text books which can provide a thorough introduction to methods, in particular, @Marescot2013 and @Smith1981.  However, implementation of these approaches generally depends on numerical methods, for which even precise mathematical formulae or pseudo-code descriptions can be inadequate to specify unambiguously or to properly facilitate replication [@Barnes2010].  As such, this appendix includes computer code in the R language, which can also be found in the corresponding R-Markdown document in the paper's GitHub repository, <https://github.com/cboettig/bad-forecast-good-decision>.  While computer code can be more verbose and difficult to read than mathematical formulae presented here, it is also less ambiguous. Though R is already widely used among ecologists, the implementations concern mostly matrix algebra that can be translated to other languages, and the `MDPtoolbox` package used here is also implemented in Matlab [@Marescot2013; @MDPtoolbox].

## Optimal quotas as a Markov Decision Process (MDP) problem

The question of optimal harvest quotas given a population growth model can be posed as a Markov Decision Process (MDP). @Marescot2013 provides an accessible overview of MDP problems and their solutions in conservation context.  @Reed1979 provides a formal proof of optimal harvest strategy for stochastic dynamics under the assumption of any concave growth function. The problem can be summarized as follows:

@@

This problem is already well studied, and it is worth noting that even under the simple utility function which assigns no value to fish that are not harvested, the optimal strategy still seeks to sustain the fish population indefinitely; as @Clark1973 shows for the deterministic function $f$ and @Reed1979 extended to the stochastic case.  As @Reed1979 proves, the optimal strategy can be characterized by a policy of "constant escapement," in which the manager adjusts the harvest effort each year in an effort to ensure the same number or biomass of fish are left in the sea each year.  The optimization problem can then be solved by dynamic programming, using the recursive Bellman equation. This technique, known as stochastic dynamic programming (SDP) is widely used in conservation decision problems [e.g. @Mangel1985, @Marescot2013].

@@

## Solving MDP via SDP

@@

Passive adaptive management using a Bayesian learning scheme still learns the wrong model. Adaptive management for MDP problems is particularly computationally intensive, and not directly supported by the `MDPtoolbox` implementation, though the principle involves only a relatively straightforward, if not particularly efficient extension to the task of solving and MDP problem for a single model shown above.  Given a set of transition probability matrices $P_i$ and a belief probability $b_i$ that the $i$th model is correct, it simple to define the transition matrices over possible models by the laws of conditional probability, $P = \sum_i P_i b_i$, given $\sum_i b_i = 1$. After each subsequent action and then new observation x_j, the $b_i$ probabilities for each model are updated according to Bayes law: if $b_i = P(M_i)$ is the prior probability that the model $M_i$ is correct, then the posterior probability that model $M_i$ is correct given observation $x_j$ is:

$$b_i' = P(M_i | x_j) = P(M_i) P(x_j | M_i) = b_i P(x_j | M_i) $$

Note that because the transition matrices are now altered after every action and observation, it is necessary to repeat the complete SDP calculation to determine the optimal action given the updated uncertainty.  While straight forward, this is what makes adaptive management for MDP problems particularly computationally intensive and thus poorly suited for handling a large space of candidate models. Nevertheless, our example of two models is most feasible, and even the larger example of 36 models (appendix B) is not unreasonable. Here I make use of the `mdplearning` R package, which provides a straight-forward implementation of this algorithm.

@@

```{r}
am <- am1$df %>% mutate(belief = am1$posterior$V1,
                        stock = states[state],
                        quota = actions[action])
write_csv(am, "../data/am.csv")

### 2021-03-26 manuscript/appendix_B.Rmd
:pen:

---
title: 'Appendix for: "Bad Forecast, Good Decision"'
date: "`r Sys.Date()`"
author: Carl Boettiger
journal: Appendix
layout: 3p
bibliography: refs.bib
#output:
#  tufte::tufte_handout:
#    dev: cairo_pdf
#    latex_engine: xelatex
output: 
  hrbrthemes::ipsum_pdf:
    dev: cairo_pdf
    latex_engine: xelatex

>>>>

---
title: 'Appendix B'
date: "`r Sys.Date()`"
author: Carl Boettiger
journal: Appendix
layout: 3p
bibliography: refs.bib
#output:
#  tufte::tufte_handout:
#    dev: cairo_pdf
#    latex_engine: xelatex
output: 
  hrbrthemes::ipsum_pdf:
    dev: cairo_pdf
    latex_engine: xelatex

### 2021-03-26 manuscript/manuscript.Rmd
:pen:

---
title: 'Are we learning to build better models at the cost of better decisions?'
author:
- name: Carl Boettiger
  email: cboettig@berkeley.edu
  affiliation: a,1
date: "`r Sys.Date()`"

@@

<!-- Open on real-world issues, uncertainty, particularly model uncertainty as key theme.  -->
Global change issues are complex and outcomes are difficult to predict  [@Clark2001]. 
To guide decisions in an uncertainty world, researchers and decision makers 
may consider a range of alternative plausible models[^1] to better reflect 
what we do and do not know about the processes involved [@Polasky2011].  Forecasts or
predictions from possible models can indicate what outcomes are most likely to result
under what decisions or actions. This has made model-based forecasts a cornerstone for scientifically
based decision making.  But, do better models always lead to better decisions?

Whenever models are estimated from the data they seek to describe, there is a concern that
the model may over-fit the data: making the data seem even more likely than it would be
under the true process [@Burnham1998].  Over-fitting is a central concern of any model comparison
because over-fit models may underestimate uncertainty and make poor predictions about future
outcomes [@Ginzburg2004]. By comparing outcomes predicted by a model to future observations, 
a decision maker can not only _plan_ for the uncertainty, but also _learn_ which models are most trustworthy.  
This iterative learning process of learning will not only decrease uncertainty by narrowing 
beliefs over plausible model or parameter values, but can also eliminate models which overfit the data.
Because any iterative learning compares models to future outcomes, overfitting those outcomes
is impossible: a concept formalized rigorously by @Gneiting2007's proof of "proper" scoring rules.
By definition, under a proper scoring rule, no probabilistic prediction $Q(x)$ can score better, on
average, than that of the underlying process, $P(x)$.  @Gneiting2007 further shows which
techniques for evaluating forecast predictions do an do not satisfy this condition, i.e. which
cannot be over-fit.  The ability to decrease uncertainty while avoiding over-fitting makes
such iterative learning particularly appealing approach to decision making in an uncertain
world [@Dietze2018], and the approach is already well established in the principle of 
adaptive management [@Walters1978] and existing practices such as Management Strategy Evaluation (MSE)
[@Punt2016] used in fisheries. Given this ability to decrease uncertainty and avoid over-fitting,
such iterative updating approaches seem irreproachable from a theoretical basis, even if
they may be difficult to execute in practice.

However, as long as the true process model is unknown (effectively always), there is no
such theoretical guarantee that iterative learning over possible models or parameters will
lead to better outcomes.  This is not -- cannot -- be because some models have over-fit the 
data, as least as long as proper scoring rules are used for evaluation.  As I illustrate here,
given a set of candidate models, iterative learning either through assessment of forecast
skill or through formal adaptive management can lead to worse outcomes than forgoing any learning step.
Put another way, the value of information [VOI; @Howard1966; @Katz1987], as measured by the expected utility given that 
information minus the utility without it, can be negative: a rational decision maker ought to pay
not to know information even when that information is perfectly accurate.  
This situation is a direct consequence of the fact that a model need only capture certain
features of the underlying process to lead to an effective decision.

[^1]: While some authors distinguish between model uncertainty and parameter uncertainty, I will refer collectively to
any uncertainty that arises from our imperfect knowledge of the system as "model uncertainty," because the distinction
is largely a consequence of notation rather something more intrinsic.  For example, two structurally different models, $f(x) = a x$
and $f(x) = a x^2$ can be considered merely different parameterizations of $f(x) = a x^b$, or vice versa.

@@

@@

A manager seeks each year to select the harvest $H_t$ which will maximize the sum
of the utility derived from such a harvest and such a state, $U(X_t,H_t)$, over all time,
subject to discount rate $\delta$ [@Clark1973]:

@@

Further we imagine that the function $f$ is not known precisely, and so we will rely on an evaluation of forecasting skill across a set of candidate models to determine which one to use to manage the fishery.  Again for simplicity, we will restrict ourselves to two simple candidate models $f_1$ and $f_2$.  Both share the same underlying structure of logistic recruitment (known as the Gordon-Schaefer model in fisheries context owing to groundbreaking work independently by @Schaefer1954 and @Gordon1954), differing only in their choice of certain parameters:

@@

Where $\xi_t(\sigma)$ represents log-normal random noise with mean of unity and log-standard-deviation $\sigma$.
Model 1 is given by $r_1 = 2$, $K_1 = 16$, $\sigma_1 = 0.05$, Model 2 by $r_2 = 0.5$,  $K_2 = 10$, $\sigma = 0.075$ (in dimensionless units).  Having both the larger growth rate and the larger carrying capacity, Model 1 is clearly the more optimistic of the two choices.

@@

<!-- I feel like parts of the below paragrph should be in the abstract (i.e. "I illustrate how the most promising techniques from statistical approaches and decision theoretic approaches to decision making under uncertainty would be applied to this simple problem, and demonstrate that it both cases they lead us away from the model that produces the most desirable decisions towards worse outcome") -->

@@

I illustrate the process of model selection by strictly proper scoring rules using two scenarios. In Scenario A (passive observation) the fish stock is unharvested and allowed to recover towards carrying capacity (as simulated under our "true" model, Model 3) while comparing the observed stock size in each subsequent time step to the distribution predicted under Model 1 and Model 2 respectively [Fig 1].  The mean, $\mu$ and variance, $\sigma$ of the forecast are compared against the true observation $x$ using a Proper scoring rule of @Gneiting2007,

\begin{equation}
-(\mu - x )^2 / \sigma^2  - \log(\sigma) \label{proper}
\end{equation}

@@

Any adaptive management strategy updates posterior distributions over model uncertainty [@Punt2016; @Ludwig1982]. Unfortunately, any such adaptive updating leads to worse outcomes than the equivalent non-adaptive strategy, in which model uncertainty is held fixed.  I illustrate the application of a passive adaptive management strategy to this simple example, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty. Passive adaptive management for a simple sequential decision problem is straight forward to implement over a discrete set of states and actions using dynamic programming with iterative updates [@Smith1981, example code in Appendix]. To demonstrate that the behavior is not driven by failure to explore sufficiently, (which might be addressed by an active adaptive management), I will assign initial probability that model 2 is true at 1%.  After a single iteration of learning, these probabilities are completely reversed, with the manager deciding that model 2 is almost certainly correct  [Fig 3]. As before, this results in a management practice with much worse ecological and economic outcomes than would have been realized by a manager who stubbornly clung to model 1 without updating [Fig 2, achieving a net present value over 100 time steps that only 31% that expected under management using Model 1 alone].

```{r figure3, fig.cap="Adaptive management under model uncertainty, color indicates the belief that Model 1 is correct (blue). Each model is assigned an initial prior belief. To ensure that any issues with passive exploration are not the cause of preferring Model 2, the initial belief in Model 2 is set to 1%.  Within a single iteration of adaptive management, the belief over models is updated to near certainty in Model 2, resulting in higher harvests and lower stock sizes similar to managing under Model 2 alone, Fig 2."}
am <- read_csv("../data/am.csv")
am %>%  filter(time <=20) %>%
  ggplot(aes(time, stock, col = belief)) + 
  geom_line() + 
  geom_point() +
  geom_point(aes(time, quota), col="darkgreen", shape=2) + 
  scale_colour_gradient(limits = c(0,1), low = pal[4], high = pal[1]) +
  ylab("stock size")
```

Increasing the space of possible models to cover a whole plausible range of parameters $r$ and $K$ does nothing to resolve this problem.  Learning is somewhat slower over the larger range of possibilities, but nevertheless converges towards parameter values with lower growth rates and lower carrying capacity, for which the target biomass is far lower [Fig 4]. The manager becomes trapped in vicious cycle as biomass declines further and further, subsequent observations seem more and more consistent with models which most favor over-fishing in the first place.  As biomass declines, annual harvests and profits also plunge, and yet our manager becomes only more confident in their decisions.

```{r}
am_multi <- read_csv("../data/am_multi.csv")
am_multi %>%
  ggplot(aes(time, stock, col = belief)) + 
  geom_line() + 
  geom_point() +
  geom_point(aes(time, harvest), col="darkgreen", shape=2) + 
  scale_colour_gradient(limits = c(0,1), low = pal[4], high = pal[1]) + 
  ylab("fish stock") + labs(caption = "Belief in model 1 falls off rapidly as magagement pushes stocks to lower and lower sizes.")
```

@@

<!-- quick summary of what was demonstrated -->
Given this simple decision problem in which one of the two models leads to effectively optimal ecological and economic outcomes, current approaches invariably choose the other. Moreover, a decision maker employing an adaptive management or iterative forecasting assessment such as those considered here would have no way of realizing that the outcomes they experienced were in fact sub-optimal.  In both approaches, the manager quickly concludes that Model 1 is entirely implausible, while finding that Model 2 is remarkably accurate at predicting future values. This problem is not addressed by re-estimating parameters (equivalently, considering a larger suite of models with all possible $K$ and $r$ values): in fact leads to a model with slightly lower $K$ and even higher level of over-fishing and under-performance than Model 2 (Fig 4).

<!-- Intuition, Figure 5 -->
The reason for Model 1's seemingly contradictory ability to make good decisions but bad forecasts becomes obvious once we compare both curves to that of the underlying model, Model 3.  Plotting the growth rate functions of each model, [Fig 5A], it is hardly surprising that no method exists which would not prefer the closely overlapping Model 2 to the no-where-close Model 1 as the better approximation of Model 3. However, Nevertheless, decisions based on Model 1 are nearly indistinguishable from those based on the true model [Fig 5B], while Model 2 leads to over-harvesting.  The explanation comes from noticing that the stock size corresponding to the maximum growth rate under Model 1 (the peak of the curve) falls at almost exactly the same stock size as that of the peak growth rate for Model 3.  Meanwhile, the peak of Model 2 occurs at a substantially lower stock size.  While the optimal control solution appears to depend only on a step-ahead forecast accuracy (indeed, the SDP solution method used here takes only step-ahead forecast probabilities as input, [@Marescot2013]), mathematical analysis showed long ago [@Reed1979] that the optimal solution for this problem depends only on keeping the biomass at the value responsible for the maximum growth rate.  This realization is quite general: for most decision problems, simple models will exist under which the optimal decision is the same as it would be for the true model, even when that simple model is wrong in most other ways.

@@

<!-- Not just "get more models" -->
This case cannot be dismissed merely as being dealt a bad hand in having to pick only between Model 1 and Model 2. Model 1 approximates the key feature, giving nearly optimal outcomes.  Previous literature has underscored the importance of Knightian "unknown unknowns" two alternative models examined here fail to fully reflect our uncertainty in the underlying dynamics, and the consequences of underestimating model uncertainty are well understood [@Polasky2011; @Wintle2010]. Had we included the true model in the set of possibilities, the techniques illustrated would have had little difficulty in distinguishing it from Models 1 and 2 after sufficient iterations. In practice, we never have the true model, and even a much more thorough suite of candidate models, such as considering any possible values of $r$, $K$ [Fig 4], or even among common alternative models with different structural forms, such as Beverton-Holt [@Beverton1957], Ricker [@Ricker1954], or Shepherd [@Shepherd1980], would fail to do any better, since all these models still have symmetric growth rates with a peak growth rate at half the steady-state population size).  Models will always be simpler than reality.  To insist that this issue can and will be avoided by always including more and better models of the process misses the point.

>>>>

---
title: 'When to ignore perfect information: how iterative learning can lead to worse decisions'
author:
- name: Carl Boettiger
  email: cboettig@berkeley.edu
  affiliation: a,1
date: "`r Sys.Date()`"

@@

<!-- Open on real-world issues, uncertainty, particularly model uncertainty as key theme.  -->
Global change issues are complex and outcomes are difficult to predict  [@Clark2001]. 
To guide decisions in an uncertainty world, researchers and decision makers 
may consider a range of alternative plausible models to better reflect 
what we do and do not know about the processes involved [@Polasky2011].  Forecasts or
predictions from possible models can indicate what outcomes are most likely to result
under what decisions or actions. This has made model-based forecasts a cornerstone 
for scientifically based decision making.  
By comparing outcomes predicted by a model to future observations, 
a decision maker can not only _plan_ for the uncertainty, but also _learn_ which models are most trustworthy.  The value of iterative learning has long been reflected in the theory of 
adaptive management [@Walters1978] as well as in actual adaptive management practices such as 
Management Strategy Evaluation (MSE) [@Punt2016] used in fisheries, and is a central tenant
of a rapidly growing interest in ecological forecasting [@Dietze2018].
But, do iterative learning approaches always lead to better decisions?

In this paper, I will demonstrate that iterative learning or iterative forecasting can also be counter-productive to management outcomes:
that in some cases, a manager would be better to ignore new information which could reduce
uncertainty over available models. Put another way, the value of information (VOI, as
measured by the expected utility given that information minus the utility without it; see @Howard1966; @Katz1987), can actually be negative. This is not a consequence of imperfect information
(e.g. observation error) or the result of model over-fitting.  The issue is an intrisic 
consequence of the fact that many models can lead to the same decisions, and no generic
solution to the problem exists. However, as this example shows, by better understanding
what aspects of model are most and least important to shaping a particular decision, 
we can at least anticipate the circumstances that make this problem more likely.

Iterative learning approaches such as adaptive management or iterative forecasting are
particularly compelling because they address a different reason why seemingly-accurate
models can lead to poor outcomes: the problem of overfitting.  When models are estimated
from the data they seek to describe, the model may over-fit the data, making the data
seem even more likely than it would be under the true process [@Burnham1998].
Over-fit models may underestimate uncertainty and make poor predictions about future
outcomes [@Ginzburg2004]. Because any iterative learning compares models to future outcomes,
overfitting those outcomes is impossible.  This concept was rigorously formalized
by @Gneiting2007's proof of "proper" scoring rules. By definition, under a proper 
scoring rule, no probabilistic prediction $Q(x)$ can score better, on
average, than that of the underlying process, $P(x)$.  @Gneiting2007 further shows which
techniques for evaluating forecast predictions do an do not satisfy this condition, i.e. which
cannot be over-fit. Adaptive management is based on equally secure footing because
model predictions are again compared to future outcomes not used to fit the model.
Instead of scoring and selecting among alternative models, adaptive management 
considers probabilities over all models [@Walters1978].  These two approaches are
particularly compelling because they can both reduce uncertainty and avoid over-fitting.
My example is not meant to undercut the importance of these approaches.  Previous
work has long acknowledged the panelope of ways in which iterative learning or other
model-based decision making can go astray due to conflicting incentives,
implementation errors, or lack of resources for monitoring and updating [e.g. @Ludwig1993].
Here, the problem is more fundamental.

@@

For illustrative purposes, we will focus on the simplest example of model uncertainty,
considering two alternative models of equal complexity but differing in the estimated
value of certain parameters. While some authors distinguish between model uncertainty and parameter uncertainty, I will refer collectively to any uncertainty that arises from our imperfect knowledge of the system as "model uncertainty," because the distinction is largely a consequence of notation rather something more intrinsic.  (For example, two structurally different models, $f(x) = a x$ and $f(x) = a x^2$ can be considered merely different parameterizations of $f(x) = a x^b$, or vice versa.)

@@

A manager seeks each year to select the harvest quota $H_t$ which will maximize the sum
of the utility derived from such a harvest and such a state, $U(X_t,H_t)$, over all time,
subject to discount rate $\delta$ [@Clark1973]:

@@

For simplicity and comparison with prior work, we will assume a fixed price of fish $p$ with no additional cost on additional harvest, $U(X_t, H_t) = p \min(H_t, X_t)$ (noting that realized harvest cannot exceed the stock size). Without loss of generality we will set the price $p = 1$ and modest discount $\delta = 0.99$.

We further imagine that the function $f$ is not known precisely.  Again for simplicity, we restrict ourselves to two simple candidate models $f_1$ and $f_2$.  Both share the same underlying structure of logistic recruitment (known as the Gordon-Schaefer model in fisheries context owing to groundbreaking work independently by @Schaefer1954 and @Gordon1954), differing only in their choice of certain parameters:

@@

For simplicity, we will assume $\xi_t(\sigma)$ represents log-normal random noise with mean of unity and log-standard-deviation $\sigma_i$ for each model. Model 1 is given by $r_1 = 2$, $K_1 = 16$, $\sigma_1 = 0.05$, Model 2 by $r_2 = 0.5$,  $K_2 = 10$, $\sigma = 0.075$ (in dimensionless units).  Having both the larger growth rate and the larger carrying capacity, Model 1 is clearly the more optimistic of the two choices.

@@

We will use this example to illustrate two alternative approaches for iterative learning over model uncertainty: iterative forecasting and adaptive management.  The central difference in the approaches
is that iterative forecasting is premised on the ability to score the predictions of alternative models.
Iterative forecasting is silent on the issue of what to do with those scores, this is left up to the 
decision-maker.  Adaptive management approaches, by contrast, explicitly seek to integrate probabilities over all candidate models to reach a decision.  I consider each in turn.

@@

I illustrate the process of model selection by strictly proper scoring rules using two scenarios. In Scenario A (passive observation) the fish stock is unharvested and allowed to recover towards carrying capacity (as simulated under our "true" model, Model 3) while comparing the observed stock size in each subsequent time step to the distribution predicted under Model 1 and Model 2 respectively [Fig 1].  The mean, $\mu_t$ and variance, $\sigma_t$ of the forecast are compared against the true observation $x_t$ using a Proper scoring rule of @Gneiting2007,

\begin{equation}
S(x_t|\mu_t,\sigma_t) = -(\mu_t - x_t )^2 / \sigma_t^2  - \log(\sigma_t) \label{proper}
\end{equation}

@@

Any adaptive management strategy updates posterior distributions over model uncertainty [@Punt2016; @Ludwig1982]. Unfortunately, any such adaptive updating leads to worse outcomes than the equivalent non-adaptive strategy, in which model uncertainty is held fixed.  I illustrate the application of a passive adaptive management strategy to this simple example, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty. Passive adaptive management for a simple sequential decision problem is straight forward to implement over a discrete set of states and actions using dynamic programming with iterative updates [@Smith1981, example code in Appendix]. To demonstrate that the behavior is not driven by failure to explore sufficiently, (which might be addressed by an active adaptive management), I will assign initial probability that model 2 is true at 1%.  After a single iteration of learning, these probabilities are completely reversed, with the manager deciding that model 2 is almost certainly correct  [Fig 3]. As before, this results in a management practice with much worse ecological and economic outcomes than would have been realized by a manager who stubbornly clung to model 1 without updating, which achieves a net present value over 100 time steps that only 31% that expected under management using Model 1 alone [Fig 2].

```{r figure3, fig.cap="Adaptive management under model uncertainty, color indicating the belief that Model 1 is correct (blue), lines indicating stock biomass over time, green triangles indicating harvest quotas set by adaptive management. Panel A: Each model is assigned an initial prior belief. To ensure that any issues with passive exploration are not the cause of preferring Model 2, the initial belief in Model 2 is set to 1%.  Within a single iteration of adaptive management, the belief over models is updated to near certainty in Model 2, resulting in higher harvests and lower stock sizes similar to managing under Model 2 alone, Fig 2A. Panel B: given 36 candidate models over a broad range of parameters, adaptive management leads to models that overharvest even more, leading to even lower stock sizes and lower harvest quotas."}
am <- read_csv("../data/am.csv")
fig3a <- am %>%  filter(time <=20) %>%
  ggplot(aes(time, stock, col = belief)) + 
  geom_line(show.legend = FALSE) + 
  geom_point(show.legend = FALSE) +
  geom_point(aes(time, quota), col="darkgreen", shape=2) + 
  scale_colour_gradient(limits = c(0,1), low = pal[4], high = pal[1]) +
  ylab("stock size") + labs(subtitle="A. Two models") + ylim(c(0,11))

am_multi <- read_csv("../data/am_multi.csv")
fig3b <- am_multi %>%
  ggplot(aes(time, stock, col = belief)) + 
  geom_line() + 
  geom_point() +  
  geom_point(aes(time, quota), col="darkgreen", shape=2) + 
  scale_colour_gradient(limits = c(0,1), low = pal[4], high = pal[1]) + 
  ylab("fish stock") + labs(subtitle="B. 36 models")+ ylim(c(0,11))

fig3a + fig3b
```

Increasing the space of possible models to cover a whole plausible range of parameters $r$ and $K$ does nothing to resolve this problem.  Learning is somewhat slower over the larger range of possibilities, but nevertheless converges towards parameter values with lower growth rates and lower carrying capacity, for which the target biomass is far lower [Fig 3B; details in appendix B]. The manager becomes trapped in vicious cycle as biomass declines further and further, subsequent observations seem more and more consistent with models which most favor over-fishing in the first place.  As biomass declines, annual harvests and profits also plunge, and yet our manager becomes only more confident in their decisions.

@@

<!-- quick summary of what was demonstrated -->
Given this simple decision problem in which one of the two models leads to effectively optimal ecological and economic outcomes, current approaches invariably choose the other. Moreover, a decision maker employing an adaptive management or iterative forecasting assessment such as those considered here would have no way of realizing that the outcomes they experienced were in fact sub-optimal.  In both approaches, the manager quickly concludes that Model 1 is entirely implausible, while finding that Model 2 is remarkably accurate at predicting future values. This problem is not addressed by re-estimating parameters (equivalently, considering a larger suite of models with all possible $K$ and $r$ values): in fact leads to a model with slightly lower $K$ and even higher level of over-fishing and under-performance than Model 2 (Fig 3B).

<!-- Intuition, Figure 4 -->
The reason for Model 1's seemingly contradictory ability to make good decisions but bad forecasts becomes obvious once we compare both curves to that of the underlying model, Model 3.  Plotting the growth rate functions of each model, [Fig 4A], it is hardly surprising that no method exists which would not prefer the closely overlapping Model 2 to the no-where-close Model 1 as the better approximation of Model 3. However, Nevertheless, decisions based on Model 1 are nearly indistinguishable from those based on the true model [Fig 4B], while Model 2 leads to over-harvesting.  The explanation comes from noticing that the stock size corresponding to the maximum growth rate under Model 1 (the peak of the curve) falls at almost exactly the same stock size as that of the peak growth rate for Model 3.  Meanwhile, the peak of Model 2 occurs at a substantially lower stock size.  While the optimal control solution appears to depend only on a step-ahead forecast accuracy (indeed, the SDP solution method used here takes only step-ahead forecast probabilities as input, [@Marescot2013]), mathematical analysis showed long ago [@Reed1979] that the optimal solution for this problem depends only on keeping the biomass at the value responsible for the maximum growth rate.  This realization is quite general: for most decision problems, simple models will exist under which the optimal decision is the same as it would be for the true model, even when that simple model is wrong in most other ways.

@@

<!-- Not just "get more models" -->
This case cannot be dismissed merely as being dealt a bad hand in having to pick only between Model 1 and Model 2. Model 1 approximates the key feature, giving nearly optimal outcomes.  Previous literature has underscored the importance of Knightian "unknown unknowns" two alternative models examined here fail to fully reflect our uncertainty in the underlying dynamics, and the consequences of underestimating model uncertainty are well understood [@Polasky2011; @Wintle2010]. Had we included the true model in the set of possibilities, the techniques illustrated would have had little difficulty in distinguishing it from Models 1 and 2 after sufficient iterations. In practice, we never have the true model, and even a much more thorough suite of candidate models, such as considering any possible values of $r$, $K$ [Fig 3B], or even among common alternative models with different structural forms, such as Beverton-Holt [@Beverton1957], Ricker [@Ricker1954], or Shepherd [@Shepherd1980], would fail to do any better, since all these models still have symmetric growth rates with a peak growth rate at half the steady-state population size).  Models will always be simpler than reality.  To insist that this issue can and will be avoided by always including more and better models of the process misses the point.
